\documentclass[11pt]{article}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage[all]{xy}		%xy-pic
\usepackage{graphics}
%\usepackage{psfrag, psfig}
\usepackage[dvips]{graphicx}


%\swapnumbers
\theoremstyle{plain}%default
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem{eg}{Example}
\newtheorem*{claim}{Claim}

\theoremstyle{remark}
\newtheorem{rmk}[thm]{Remark}
\newtheorem{notn}[thm]{Notation}

\numberwithin{equation}{section}


\title{On Tiling the Unit Square}
\author{Benson Guo and Peter Hazard}
\date{\today}

\begin{document}

\pagestyle{empty}
\noindent
Dear Benson,
\vskip 10pt
Here is a template for a Latex file to keep as a log of our project.
On the second page I have added a comments section, where we can put comments, rough work, calculations, questions, etc.
These should be added as a list, with the authors name (BG or PH resp.) and the date of the comment.
\vskip 50pt

{\flushright --PeH, \today.--}
%\flushleft
\newpage

\paragraph{Comments.}
\begin{itemize}
\item[PH -- 2015/02/01]
The sections ``Cross Sections'' and onwards are rough work, and will eventually (probably) be scrapped.
\end{itemize}
\begin{itemize}
\item[BG -- 2015/02/07]
We started the investigation of the viability of a greedy algorithm.
The algorithm works as follows:
\begin{enumerate}
\item Assume the longer side of a rectangle $R$ has a length $\frac{a}{b}$, and shorter side with a length $\frac{a\prime}{b\prime}$
\item Fill $R$ with a piece with a side length $\frac{1}{n_{0}}\times\frac{1}{n_{0}+1}  $ where $\frac{1}{n_0}$ satisfies $\frac{1}{n_0}\le\frac{a\prime}{b\prime}<\frac{1}{n_{0}-1}$
\item Cut the remaining area into 2 rectangles:
\begin{quote}
$R_0: \frac{1}{n+1}\times(\frac{a\prime}{b\prime}-\frac{1}{n_0})  $
\end{quote} 
\begin{quote}
$R_1: (\frac{a}{b}-\frac{1}{n+1})\times \frac{a\prime}{b\prime} $
\end{quote} 
\item Apply Step 1 to rectangle $R_0$.
\item Stop when $\frac{a\prime}{b\prime} - \sum \frac{1}{n_i}=0$
\item Apply Step 1 to next left-most rectangle.
\end{enumerate}
Studying  $R_0: \frac{1}{n_0+1}\times(\frac{a\prime}{b\prime}-\frac{1}{n_0})  $, we ask ourselves, is $\frac{a\prime}{b\prime}-\frac{1}{n_0}>\frac{1}{n_0+1}$?
By hypothesis, $\frac{1}{n_0}\le\frac{a\prime}{b\prime}<\frac{1}{n_{0}-1} \Rightarrow 0\le\frac{a\prime}{b\prime}-\frac{1}{n_0}<\frac{1}{n_0-1}-\frac{1}{n_0}$

$\frac{1}{n_0-1}-\frac{1}{n_0}=\frac{n_0-(n_0-1)}{n_0(n_0-1)}=\frac{1}{n_0(n_0-1)}$
Is $\frac{1}{n_0(n_0-1)}<\frac{1}{n_0+1}$?

$n_0+1<n_0^2-n_0\Rightarrow n_0^{2}-2n_0-1>0$

Hence, $\frac{a\prime}{b\prime}-\frac{1}{n_i}<\frac{1}{n_i(n_i-1)}<\frac{1}{n_i+1}$ for $n_i\ge2$.
This implicates that the shorter side will be on the vertical edges, which will approach $0$ as $\frac{1}{n_i}$ decreases. When it reaches $0$, the left side will be completely covered with smaller rectangles. 
\end{itemize}
\begin{itemize}
\item[BG -- 2015/02/16]
I coded a rough simulation of the greedy algorithm to answer the question of whether or not some rectangle will be required to be used twice. Very quickly I found a case in which the same rectangle is called upon consecutively. Take the case when you have a rectangle with dimensions $\frac{1}{n}\times\frac{1}{n+x}$. The algorithm will tile it with $\frac{1}{n+x}\times\frac{1}{n+x+1}$ and the remaining area will be, $\frac{1}{n}\times(\frac{1}{n}-\frac{1}{n+x+1})$. In this case $\frac{1}{n+x}\times\frac{1}{n+x+1}$ will be used again, provided $\frac{1}{n+x+1}$ is still the shorter side. Specifically, one case I came across was $\frac{1}{11}\times\frac{1}{140}$ which would be tiled with $\frac{1}{140}\times\frac{1}{141}$ leaving a rectangle of $\frac{130}{1551}\times\frac{1}{140}$. I will play around with the algorithm to see how we can go around this.
\end {itemize}


\newpage

\maketitle

\begin{abstract}
We consider the problem, originally due to L.\ Moser, of tiling the unit square with rectangles of side lengths $\frac{1}{n}\times\frac{1}{n+1}$ for $n=1,2,\ldots$.
This is the result of a University of Toronto Mentoring Program 2015.
\end{abstract}
\section{Introduction}
We consider the following:
\begin{quote}
{\it Problem:}
Can the unit square $[0,1]^2$ be tiled with rectangles of sides $\frac{1}{n}\times\frac{1}{n+1}$ for $n=1,2,\ldots$
\end{quote}
A more general formulation is the following:
\begin{quote}
{\it Problem:}
Find the smallest square $[0,1+\epsilon]^2$, $\epsilon\geq 0$, into which the rectangles of sides $\frac{1}{n}\times\frac{1}{n+1}$, $n=1,2,\ldots$, can be packed.
\end{quote}
According to~\cite[Section D11]{CFG94} these problems are originally due to L.\ Moser.
The best result for the second problem, to the authors knowledge, is due to Balint~\cite{Balint92, Balint98}. See also Paulhus~\cite{Paulhus98}.

\section{Preliminaries}
Let $T_0=[0,1]^2$ denote the unit square.
For each $n\in\mathbb{N}$, let $T_n$ denote the tile of side lengths $\frac{1}{n}\times\frac{1}{n+1}$.
In our considerations we shall make the following hypotheses about any tiling of $T_0$:
\begin{quote}
{\it Hypothesis 1:} 
The tiles are added parallel to the sides of $T_0$.
\end{quote}
\begin{quote}
{\it Hypothesis 2:}
For each $n\in\mathbb{N}$, the free space $T_0\setminus \bigcup_{n=1}^{n}T_i$ is connected and simply connected.
\end{quote}
Assume a tiling is given.
Let 
\begin{equation}
V_0=T_0, \qquad V_n=V_0\setminus \bigcup_{i=1}^{n}T_i, \quad i=1,2,\ldots
\end{equation}
Then, by the above hypotheses, for each positive integer $n$ the region $V_n$ is a polygon with sides aligned with the horizontal and vertical directions.
Let $V_n$ have vertices $v_n^1,\ldots,v_n^{r_n}$.
Consider the set of all horizontal and vertical lines through the $v_n^j$.
Take their complement in $V_n$.
This complement consists of finitely many rectangles $F_n^1,\ldots,F_n^{k_n}$ with sides parallel to the sides of $T_0$.
Furthermore, $V_n=\bigcup_{i=1}^{k_n} F_n^i$.

We make the following assumptions:
\begin{quote}
{\it Assumption 1:}
For each positive integer $n$, the tile $T_n$ is positioned in $V_{n-1}$ so that at least one vertex of $T_n$ lies at one of the vertices of $V_{n-1}$.
\end{quote}
\begin{quote}
{\it Assumption 2:}
For each positive integer $n$, if the tile $T_n$ fits into one of the free spaces $F_{n-1}^j$, then it must be positioned in one of them.
\end{quote}
Now we ask the following questions:
\begin{enumerate}
\item Does a tiling exist satisfying Assumptions 1 and 2?
\item Does such a tiling have the property that for each positive integer $n$, each $F_n^i$ except one, is eventually tiled by finitely many tiles?
\end{enumerate}
Since each $F_n^i$ has side-lengths of the form $\sum \frac{1}{n_i}-\sum \frac{1}{m_j}$ we therefore ask, firstly, if the sides can be `tiled' by a finite number of intervals of lengths $\frac{1}{l_1},\frac{1}{l_2},\ldots,\frac{1}{l_k}$, where $l_1,l_2,\ldots,l_k$ are distinct positive integers and $\min_r l_r >\max (n_i,m_j)$.
In the case when we do not have the second contraint we know the following, originally due to Fibonacci and rediscovered by Sylvester.
\begin{thm}
Any positive rational number $\frac{a}{b}$ can be expressed as a finite sum of unit fractions, i.e. there exist distinct positive integers $n_1,n_2,\ldots,n_k$ such that
\begin{equation}
\frac{a}{b}=\frac{1}{n_1}+\frac{1}{n_2}+\cdots+\frac{1}{n_k}
\end{equation}
\end{thm}


\subsection{Cross Sections}
Assume a tiling exists. 
Take a line parallel to the horizontal or vertical sides of $T_0$ and intersect it with $T_0$.
Assume that the tiles $T_{i_1},T_{i_2},\ldots$ intersect this line horizontally and the tiles $T_{j_1},T_{j_2}$ intersect it vertically.
Then
\begin{equation}
\sum\frac{1}{i_k}+\sum\frac{1}{j_k-1}=1
\end{equation}
The problem of finding positive integers $i_k, j_k$ satisfying this equation is classical. 
(See Klee and Wagon~\cite[Chapter 2.15]{KleeWagon91}.)

Assume there does exist a tiling of the unit square. 
Note that if we take a horizontal section, i.e. intersect with $\{y=y_0\}$ we find the length decomposes as
\begin{equation}
1=|l(S(y_0))|=\sum |l(S_{n_i(y_0)}(y_0))|=\sum\frac{1}{n_i(y_0)}
\end{equation}

\section{Perimeter Representation}
Observe that if a tiling exists then the perimeter of the free space must tend to zero, i.e. $\lim_{n\to \infty}|P(V_n)|=0$.

\subsection{Winding Numbers: How to tell adding a tile is admissible.}

\newpage
\begin{appendix}

\end{appendix}

\small
\addcontentsline{toc}{section}{Bibliography}
\bibliographystyle{amsplain}
\bibliography{BG-PH_Tiling01}


\end{document}


\begin{thebibliography}{99}
\bibitem{HFG1} Hallcroft, Kenneth Falconer and Richard Guy, ``Unsolved Problems...''.
\bibitem{KW1} Victor Klee and Stan Wagon, ``...''
\end{thebibliography}
